Familial consequences of the Covid-19 in Québec: Comparison between immigrants and natives

### 2023-05-02 Codes/Trajipaq_covid_an1.Rmd

<!--
Ce fichier est pour les statistiques descriptives. 
La base de données provient de Trajipaq_covid_cr1
Ce fichier est une extension de Trajipaq_covid_an2 où le statut d'immigration est remplacé
par une combinaison de statut et de minorité visible.

Couleur palette graphique : https://coolors.co/b224e5-f489a1-ffe385-f6fdf4-6c885e-b0c4c3-9cc8ac-afecb1-60ade1-95d3ff
-->

@@

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(lubridate)

trajipaq_covid <- read_rds("../Data/trajipaq_covid.RData")

freq(trajipaq_covid$imm_minorite)

trajipaq_covid <-
  trajipaq_covid %>% 
  filter(imm_minorite != "Others")

freq(trajipaq_covid$imm_minorite)

@@

freq(trajipaq_covid$minorite_vis)

freq(trajipaq_covid$statut_imm, weights = trajipaq_covid$Poidstot1)
freq(trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1)

ctable(trajipaq_covid$statut_imm, trajipaq_covid$minorite_vis, weights = trajipaq_covid$Poidstot1, "t")

@@

```{r}
cbPalette <- c("#56B4E9", "#E69F00", "#999999", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Fonction
graph_traj <- function(resultat){
  ggplot(resultat) +
  geom_col(aes(x = reorder(Variables, desc(Proportion)), y = Proportion, fill = imm_minorite), position = "dodge") +
  #geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2, position=position_dodge(.9))
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_flip() +
    labs(x = "", fill = "") +
  scale_fill_manual(values = cbPalette)
}

graph_traj2 <- function(resultat){
  ggplot(resultat) +
    geom_col(aes(x = Variables, y = Proportion, 
                 fill = fct_relevel(Modalités, "Much degraded",  "A little degraded",
                                    "No change", "A little improved", "Much improved"),
                 position = "fill")) +
    facet_wrap(~imm_minorite) +
#    scale_x_discrete("Modalités", limits = c("Much improved", "A little improved",
#    "No change", "A little degraded", "Much degraded")) +
    theme_bw() +
    theme(legend.position = "bottom", legend.box="vertical", legend.margin=margin()) +
    guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
#    theme(strip.background = element_rect(colour = "black", fill = "blue")) +
    coord_flip() +
    labs(x = "", fill = "") +
    #scale_fill_manual(values = cbPalette) + 
    scale_fill_brewer(palette="PRGn")
}

#gg+guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

exposition <- "e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo"
exposition
exposition[1]
noquote(exposition)[1]
class(exposition)

@@

traj_exposition <- svyby(~ e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede + expo, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_exposition

traj_exposition <-
  traj_exposition %>% 
  pivot_longer(col=!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)

@@

graph_traj_exposition <- graph_traj(traj_exposition)

ggsave(graph_traj_exposition, file = "Figures/Exposition1.png")
#ggsave(graph_traj_exposition, file = "Figures/Exposition2.png")

graph_traj_exposition
```

@@

- On getting ressources

- We had to leave our home
- We had difficulty obtaining food
- We had trouble getting medicine
- We had difficulty getting health care when we needed it
**- Our family income has decreased**
- We lost health insurance / benefits

@@

# f_separe + f_emmenage + f_revenu +

cons_famille <- "f_quitte_dom + f_nourriture + f_medicaments + f_sante +  f_assurance + f_evenement + f_essentiel"

#f_quitte_dom +  f_separe + f_emmenage + + f_revenu + f_evenement

traj_famille <- svyby(~f_nourriture + f_medicaments + f_sante + f_assurance + f_essentiel, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_famille

traj_famille <-
  traj_famille %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)
traj_famille

traj_famille <-
  traj_famille %>% 
  mutate(Variables = case_when(
    #Variables == "f_separeoui" ~ "Live separed",
    #Variables == "f_emmenageoui" ~ "Someone moved to our home",
    #Variables == "f_quitte_domoui" ~ "Leave our home",
    Variables == "f_nourritureoui" ~ "Difficulty obtaining food",
    Variables == "f_medicamentsoui" ~ "Difficulty getting medicine",
    Variables == "f_santeoui" ~ "Difficulty getting health care",
    #Variables == "f_revenuoui" ~ "Family income has decreased",
    Variables == "f_assuranceoui" ~ "Lost family insurrance",
#    Variables == "f_evenementoui" ~ "Missed important family event",
    Variables == "f_essentieloui" ~ "difficulty obtaining other essentials"))

@@

graph_traj_famille <- graph_traj(traj_famille)

ggsave(graph_traj_famille, file = "Figures/Famille1.png")
graph_traj_famille

@@

### Conséquences emploi et revenu

@@

traj_emploi <- svyby(~ f_reduire + f_cesser + f_perdu + f_revenu, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_emploi

traj_emploi <-
  traj_emploi %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)

@@

traj_emploi <-
  traj_emploi %>% 
  mutate(Variables = case_when(
    Variables == "f_reduireoui" ~ "Has reduced working hours",
    Variables == "f_cesseroui" ~ "Has to stop working",
    Variables == "f_perduoui" ~ "Has lost their job permanently",
    Variables == "f_revenuoui" ~ "Family income has decreased"))

#A family member

@@

ggsave(graph_traj_emploi, file = "Figures/Emploi1.png")

@@

*On ability to perform works*

COVID-19 can have many impacts on you and your family life. In general, how has the COVID-19 pandemic affected each of the following?

- Our family lived separately for health, safety or employment reasons
- How family members get along
- Ability to care for your child with an illness / condition
- Ability to take care of other children in your family
- Ability to care for the elderly or disabled in your family
- We missed an important family event or it was canceled (for example, wedding, graduation, birth, funeral, travel (including vacations), other

@@

traj_entente <- svyby(~ f_separe + f_role_parent + f_entende + f_enfant_maladie + f_autre_enfant + f_pers_agees + f_evenement, ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_entente

traj_entente <-
  traj_entente %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_entente

@@

traj_entente <-
  traj_entente %>% 
  filter(Modalités != "Not applicable") %>% 
  mutate(Variables = case_when(
    Variables == "f_role_parent" ~ "Parenting",
    Variables == "f_separeoui" ~ "Live separed",
    Variables == "f_entende" ~ "How family members get along",
    Variables == "f_enfant_maladie" ~ "Ability to care for your child with an illness",
    Variables == "f_autre_enfant" ~ "Ability to take care of other children",
    Variables == "f_pers_agees" ~ "Ability to care for the elderly or disabled",
    Variables == "f_evenementoui" ~ "Missed important family event"))

@@

ggsave(graph_traj_entente, file = "Figures/Entente1.png")

@@

traj_bienetre <- svyby(~ be_exercice + be_manger + be_dormir + be_anxiete + be_humeur, 
                       ~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_bienetre

traj_bienetre <-
  traj_bienetre %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_bienetre

@@

ggsave(graph_traj_bienetre, file = "Figures/Bien-être1.png")

@@

traj_aide <- svyby(~ patt + pcu,~ imm_minorite, trajipaq.design, svymean, na.rm = TRUE)
traj_aide

traj_aide <-
  traj_aide %>% 
  pivot_longer(!imm_minorite, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "Oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)
traj_aide

@@

ggsave(graph_traj_aide, file = "Figures/Aide1.png")

@@

## Covid data from statistics canada

covid_canada <- read_csv("../Data/covid19_Canada_281121.csv")

quebec <- covid_canada %>% 
  filter(prname == "Quebec") %>% 
  mutate(Period = format_ISO8601(date, precision = "ym")) %>% 
  group_by(Period) %>% 
  summarise('Number of confirmed cases' = sum(numconf))

Quebec_cases <- ggplot(quebec) +
  geom_line(aes(x = Period, y = `Number of confirmed cases`, group = 1), color = "blue") +
  geom_vline(xintercept = "2020-08", linetype = "twodash") +
  geom_vline(xintercept = "2020-11", linetype = "twodash") +
#  geom_ribbon(aes(ymin = 0, ymax = 200000, xmin = "2020-08", xmax = "2020-11")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
Quebec_cases

ggsave(Quebec_cases, file = "Figures/Quebec_cases1.png")

>>>>

<!--
Ce fichier est pour les statistiques descriptives. 
La base de données provient de Trajipaq_covid_cr1

-->

@@

library(tidyverse)
library(summarytools)
library(srvyr)
library(survey)
library(ggpubr)
library(writexl)
library(haven)

Base_Trajipaq_Finale_Ok <- read_dta("~/GitHub/Projet_Immigration_Trajipaq/Data/Base_Trajipaq_Finale_Ok.dta")

@@

@@

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Fonction
graph_traj <- function(resultat){
  ggplot(resultat) +
  geom_col(aes(x = reorder(Variables, desc(Proportion)), y = Proportion, fill = statut_imm), position = "dodge") +
  #geom_errorbar(aes(ymin=len-sd, ymax=len+sd), width=.2, position=position_dodge(.9))
  theme_bw() +
  theme(legend.position = "bottom") +
  coord_flip() +
    labs(x = "", fill = "") +
  scale_fill_manual(values = cbPalette)
}

graph_traj2 <- function(resultat){
  ggplot(resultat) +
    geom_col(aes(x = Variables, y = Proportion, 
                 fill = fct_relevel(Modalités, "Much degraded",  "A little degraded",
                                    "No change", "A little improved", "Much improved"),
                 position = "fill")) +
    facet_wrap(~statut_imm) +
#    scale_x_discrete("Modalités", limits = c("Much improved", "A little improved",
#    "No change", "A little degraded", "Much degraded")) +
    theme_bw() +
    theme(legend.position = "bottom", legend.box="vertical", legend.margin=margin()) +
#    theme(strip.background = element_rect(colour = "black", fill = "blue")) +
    coord_flip() +
    labs(x = "", fill = "") +
    #scale_fill_manual(values = cbPalette) + 
    scale_fill_brewer(palette="PRGn")
}

@@

exposition <- "e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede"
exposition
exposition[1]
noquote(exposition)[1]
class(exposition)

@@

traj_exposition <- svyby(~ e_hors_dom + e_four_soin + e_quarantaine + e_expose + e_symptome + e_hospitalise + e_intensif + e_decede, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_exposition

traj_exposition <-
  traj_exposition %>% 
  pivot_longer(col=!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)

@@

graph_traj_exposition <- graph_traj(traj_exposition)
graph_traj_exposition

ggsave(graph_traj_exposition, file = "Figures/Exposition.png")

@@

@@

cons_famille <- "f_separe + f_emmenage + f_quitte_dom + f_nourriture + f_medicaments + f_sante + f_revenu + f_assurance + f_evenement + f_essentiel"

traj_famille <- svyby(~ f_separe + f_emmenage + f_quitte_dom + f_nourriture + f_medicaments + f_sante + f_revenu + f_assurance + f_evenement + f_essentiel, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_famille

traj_famille <-
  traj_famille %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_detect(Variables, "se")) %>% 
  arrange(Variables)
traj_famille

traj_famille <-
  traj_famille %>% 
  mutate(Variables = case_when(
    Variables == "f_separeoui" ~ "Live separed",
    Variables == "f_emmenageoui" ~ "Someone moved to our home",
    Variables == "f_quitte_domoui" ~ "Leave our home",
    Variables == "f_nourritureoui" ~ "Difficulty obtaining food",
    Variables == "f_medicamentsoui" ~ "Difficulty getting medicine",
    Variables == "f_santeoui" ~ "Difficulty getting health care",
    Variables == "f_revenuoui" ~ "Family income has decreased",
    Variables == "f_assuranceoui" ~ "Lost family insurrance",
    Variables == "f_evenementoui" ~ "Missed important family event",
    Variables == "f_essentieloui" ~ "difficulty obtaining other essentials"))

@@

graph_traj_famille <- graph_traj(traj_famille)
graph_traj_famille

ggsave(graph_traj_famille, file = "Figures/Famille.png")

@@

### Conséquences emploi

@@

traj_emploi <- svyby(~ f_reduire + f_cesser + f_perdu, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_emploi

traj_emploi <-
  traj_emploi %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)

@@

traj_emploi <-
  traj_emploi %>% 
  mutate(Variables = case_when(
    Variables == "f_reduireoui" ~ "Has reduced working hours",
    Variables == "f_cesseroui" ~ "Has to stop working",
    Variables == "f_perduoui" ~ "Has lost their job permanently"))
#A family member

@@

ggsave(graph_traj_emploi, file = "Figures/Emploi.png")

@@

@@

traj_entente <- svyby(~ f_role_parent + f_entende + f_enfant_maladie + f_autre_enfant + f_pers_agees, ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_entente

traj_entente <-
  traj_entente %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_entente

@@

traj_entente <-
  traj_entente %>% 
  filter(Modalités != "Not applicable") %>% 
  mutate(Variables = case_when(
    Variables == "f_role_parent" ~ "Parenting",
    Variables == "f_entende" ~ "How family members get along",
    Variables == "f_enfant_maladie" ~ "Ability to care for your child with an illness",
    Variables == "f_autre_enfant" ~ "Ability to take care of other children",
    Variables == "f_pers_agees" ~ "Ability to care for the elderly or disabled"))

@@

ggsave(graph_traj_entente, file = "Figures/Entente.png")

@@

traj_bienetre <- svyby(~ be_exercice + be_manger + be_dormir + be_anxiete + be_humeur, 
                       ~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_bienetre

traj_bienetre <-
  traj_bienetre %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(!str_starts(Variables, "se")) %>% 
#  separate(Variables, into = c("Variables", "Modalités"), sep = "[[:upper:]]") %>% 
  separate(Variables, into = c("Variables", "Modalités"), sep = "(?=[A-Z])") %>% 
  arrange(Variables)
traj_bienetre

@@

ggsave(graph_traj_bienetre, file = "Figures/Bien-être.png")

@@

traj_aide <- svyby(~ patt + pcu,~ statut_imm, trajipaq.design, svymean, na.rm = TRUE)
traj_aide

traj_aide <-
  traj_aide %>% 
  pivot_longer(!statut_imm, names_to = "Variables", values_to = "Proportion") %>% 
  filter(str_detect(Variables, "Oui") & !str_starts(Variables, "se")) %>% 
#  replace(Variables, values = c("f_cesseroui", "f_perdreoui", "f_reduireoui"), list = c("Cesser de travailler", "Perdre son emploi", "Réduire le temps de travail"))
  arrange(Variables)
traj_aide

@@

ggsave(graph_traj_aide, file = "Figures/Aide.png")

@@

## Modèle de régressions
